
---
title: "P8105 Homework 2"
author: "Tianqing Wang"
output: github_document
---

```{r}
library(tidyverse)
```

# Problem 1

```{r}
subway_df = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
col_types = cols(
  Route8 = col_character(),
  Route9 = col_character(),
  Route10 = col_character(),
  Route11 = col_character()
)
)
  
subway_df = janitor::clean_names (subway_df)
names(subway_df)
head(subway_df)
```


```{r}
subway_df = subway_df |>
  select(
    line,
    station_name,
    station_latitude,
    station_longitude,
    starts_with("route"),
    entry,
    vending,
    entrance_type,
    ada
  )
dim(subway_df)
names(subway_df)
```


```{r}
subway_df = subway_df |>
  mutate(
    entry = case_match(
    entry,
    "YES" ~ TRUE,
    "NO" ~ FALSE
  )
  )
  
subway_df |>
  select(entry) |>
  head()
```


### Dataset Description
The dataset contains information about NYC subway station entrances and exits, including subway lines, station names, station latitude and longitude, routes served, entry permission, vending availability, entrance types, and ADA accessibility.

I cleaned the data by standardizing variable names using janitor::clean_names(), selecting the relevant variables using select(), and converting the entry variable from character ("YES" and "NO") to logical (TRUE and FALSE) using mutate() and case_match().

The resulting dataset contains 1,868 rows and 19 columns.

The data are not fully tidy because route information is spread across multiple columns (route1 to route11) rather than being organized into separate variables for route number and route name.


### Distinct Stations
```{r}
subway_df |>
  distinct (line, station_name) |>
  dim ()

subway_df |>
  filter (ada == TRUE) |>
  distinct (line, station_name) |>
  dim ()
```
There are 465 distinct subway stations in the dataset, identified by unique combinations of station name and line. Among these, 84 stations are ADA-compliant.


### Vending and Entry
```{r}
subway_df |>
  filter (vending == "NO") |>
  pull(entry) |>
  mean()
```
Among subway station entrances and exits without vending facilities, approximately 37.7% allow entry.

### Subway Routes
```{r}
subway_long_df = subway_df |>
  pivot_longer(
    route1 : route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name"
  ) |>
drop_na(route_name)
head(subway_long_df)

subway_long_df |>
  select(station_name, route_number, route_name) |>
  head()
```

### Distinct Stations Served by the A Train
```{r}
subway_long_df |>
  filter (route_name =="A") |>
  distinct (line, station_name) |>
  dim()
```
There are 60 distinct subway stations served by the A train, identified by unique combinations of station name and line.

### ADA-Compliant Stations Served by the A Train
```{r}
subway_long_df |>
  filter(route_name == "A", ada == TRUE) |>
  distinct(line, station_name) |>
  dim()
```
Among the 60 distinct stations served by the A train, 17 are ADA-compliant.




#######################################  Problem 2  ################################################



# Problem 2
```{r}
library(readxl)
```

################################## Mr. Trash Wheel ################################################


## Import Mr. Trash Wheel data
```{r}
mr_trash_df=read_excel(
  "./data/202610 Trash Wheel Collection Data.xlsx",
  sheet = "Mr. Trash Wheel"
)
names(mr_trash_df)
head(mr_trash_df)
tail(mr_trash_df)
dim(mr_trash_df)
```

## use reasonable variable names
```{r}
mr_trash_df = janitor::clean_names (mr_trash_df)
names(mr_trash_df)
```

## Remove unnecessary columns
```{r}
mr_trash_df = mr_trash_df |>
  select(-x15, -x16)
names(mr_trash_df)
```

## # Identify rows with missing dumpster IDs
```{r}
mr_trash_df |>
  filter (is.na(dumpster))
dim(mr_trash_df)
```

# Omit rows without dumpster-specific data
```{r}
filter(mr_trash_df, is.na(dumpster))
mr_trash_df = mr_trash_df |>
  drop_na(dumpster)
dim(mr_trash_df)
```

## Inspect sports balls before rounding
```{r}
mr_trash_df |>
  select (sports_balls)
dim(mr_trash_df)
```

## Round sports balls and convert to integer
```{r}
mr_trash_df = mr_trash_df |>
  mutate(sports_balls=as.integer(round(sports_balls)))

  mr_trash_df |>
  select (sports_balls)
dim(mr_trash_df)

```


#####################################  Professor Trash Wheel  ######################################


## Import Professor Trash Wheel data
```{r}
prof_trash_df=read_excel(
  "./data/202610 Trash Wheel Collection Data.xlsx",
  sheet = "Professor Trash Wheel"
)
names(prof_trash_df)
head(prof_trash_df)
tail(prof_trash_df)
dim(prof_trash_df)
```

# Import Professor Trash Wheel data (-diva in column 135)
```{r}
prof_trash_df = read_excel(
  "./data/202610 Trash Wheel Collection Data.xlsx",
  sheet = "Professor Trash Wheel",
  na = c("", "dive")
)
```


## use reasonable variable names
```{r}
prof_trash_df = janitor::clean_names (prof_trash_df)
names(prof_trash_df)
```

# Check dumpster 135 transform to NA
```{r}
prof_trash_df |>
  filter(dumpster == 135) |>
  select(dumpster, plastic_bottles)
```

## Identify rows with missing dumpster IDs
```{r}
prof_trash_df |>
  filter (is.na(dumpster))
dim(prof_trash_df)
```

## Omit rows without dumpster-specific data
```{r}
filter(prof_trash_df, is.na(dumpster))
prof_trash_df = prof_trash_df |>
  drop_na(dumpster)
dim(prof_trash_df)
```

########################  Gwynnda the Good Wheel of the W  #########################################

## Import Gwynnda the Good Wheel of the W
```{r}
gwynnda_df=read_excel(
  "./data/202610 Trash Wheel Collection Data.xlsx",
  sheet = "Gwynnda the Good Wheel of the W"
)
names(gwynnda_df)
head(gwynnda_df)
tail(gwynnda_df)
dim(gwynnda_df)
```

## use reasonable variable names
```{r}
gwynnda_df = janitor::clean_names (gwynnda_df)
names(gwynnda_df)
```

## Identify rows with missing dumpster IDs
```{r}
gwynnda_df |>
  filter (is.na(dumpster))
dim(gwynnda_df)
```

## Omit rows without dumpster-specific data
```{r}
filter(gwynnda_df, is.na(dumpster))
gwynnda_df = gwynnda_df |>
  drop_na(dumpster)
dim(gwynnda_df)
```

## Standardize year data type before merging in "mr_trash_df" data
```{r}
mr_trash_df = mr_trash_df |>
  mutate(year = as.integer(year))
```

############################################  Combining  ###########################################

## Add Trash Wheel identifiers
```{r}
mr_trash_df=mr_trash_df |>
  mutate(trash_wheel="Mr")
prof_trash_df=prof_trash_df |>
  mutate(trash_wheel="Prof")
gwynnda_df=gwynnda_df |>
  mutate(trash_wheel="gwynnda")
```

## Combine the three Trash Wheel datasets
```{r}
trash_wheel_df=bind_rows(
  mr_trash_df,
  prof_trash_df,
  gwynnda_df
)
dim(trash_wheel_df)
head(trash_wheel_df)
```

##Total weight collected by Professor Trash Wheel
```{r}
prof_total_weight = trash_wheel_df |>
  filter(trash_wheel == "Prof") |>
  pull(weight_tons) |>
  sum()
```

## Total number of cigarette butts collected by Gwynnda in June of 2022
```{r}
gwynnda_june_butts = trash_wheel_df |>
  filter(
    trash_wheel == "gwynnda", 
    year == 2022, 
    month =="June") |>
  pull(cigarette_butts) |>
  sum()
```

## Trash Wheel Data Summary
The combined Trash Wheel dataset contains `r dim(trash_wheel_df)[1]` observations and `r dim(trash_wheel_df)[2]` variables, including data from Mr. Trash Wheel, Professor Trash Wheel, and Gwynnda. Each observation represents a dumpster collection record. Key variables include the Trash Wheel identifier, collection date, weight of trash in tons, volume in cubic yards, and counts of different types of trash, such as plastic bottles and cigarette butts.

The data were cleaned by standardizing variable names, removing non-dumpster-specific summary rows, converting variable types where necessary, and combining the three datasets. Based on the available data, Professor Trash Wheel collected a total of `r prof_total_weight` tons of trash, while Gwynnda collected `r gwynnda_june_butts` cigarette butts in June 2022.


#######################################  Problem 3  ################################################

Import, clean, and tidy the dataset of baseline demographics. Ensure that sex and APOE4 carrier status are appropriate encoded (i.e. not numeric), and remove any participants who do not meet the stated inclusion criteria (i.e. no MCI at baseline). 

## Import and check the baseline data
```{r}
baseline_df = read_csv("./data/MCI_baseline.csv")
 names(baseline_df)
 str(baseline_df)
 dim(baseline_df)
 head(baseline_df)
 tail(baseline_df)
```

## Re-import the baseline dataset
```{r}
baseline_df = read_csv(
  "./data/MCI_baseline.csv",
  skip = 1,
  na=c ("",".")
)
   names(baseline_df)
 str(baseline_df)
 dim(baseline_df)
 head(baseline_df)
  
```

## Use reasonable variable names
```{r}
baseline_df = janitor::clean_names (baseline_df)
names(baseline_df)
head(baseline_df)
```

## Identify rows with missing age_at_onset
```{r}
baseline_df |>
  filter (is.na(age_at_onset)) 
dim(baseline_df)
```

## Remove participants without MCI at baseline
```{r}
baseline_df = baseline_df |>
filter(is.na(age_at_onset) | age_at_onset > current_age)
  dim(baseline_df)
  str(baseline_df)
```

## Recode the variables：apoe4 and sex
```{r}
baseline_df = baseline_df |>
  mutate(
    sex=case_match(
      sex,
      1 ~ "Male",
      0 ~ "Female"
    )
    ,
    apoe4=case_match(
      apoe4,
      1 ~ "APOE4 carrier", 
      0 ~ "APOE4 non-carrier"
    )
  )
str(baseline_df)
```


















